\chapter{External Interfaces}

This chapter explains the most common ways that we interface Drake
with real robots - using simple network protocols.  Typically one or
more instances of (desktop) MATLAB are up running methods/systems in Drake as
nodes in a planning/estimation/control network.

\section{Lightweight Communications and Marshalling (LCM)}

LCM is a simple, elegant, and efficient network protocol developed at
MIT during the DARPA Urban Challenge
(\url{https://code.google.com/p/lcm/}) \cite{Huang10}.  For Drake, we
primarily interact with LCM via the MATLAB/Java interface and
lcm-java.  For some high-performance applications, we do replace it
with custom C mex implementations.

\subsection{LCM Coders}

LCM passes messages defined in an lcmtype file which can describe an
almost arbitrary data structure.  Drake traffics primarily in signals,
which are (timestamped) vectors of doubles.  To convert between the two, we define
an \mcode{LCMCoder} class which defines how to \mcode{encode} a double
as the rich data structure and \mcode{decode} the data structure back
into a double.

You can implement your \mcode{LCMCoder} directly in matlab or in java.
The \mcode{LCMCoderFromType} class can also construct one
automatically using java reflection -- it simply concatenates the
elements of the structure into a double vector in the order that they
appear in the lcmtype file.  
 
\subsection{LCM Coordinate Frames}

Any coordinate frame can be associated with an \mcode{LCMCoder} by deriving
from the \mcode{LCMCoordinateFrame} class. 

In order to use any \mcode{DynamicalSystem} as a node in the LCM
network, simply call \mcode{runLCM(sys,...)} instead of
\mcode{simulate(sys,...)} -- this will automatically attach LCM input blocks
and LCM output blocks on any input/output frames of the system that
are of type \mcode{LCMCoordinateFrame}, and then simulate the system.
The \mcode{runLCM} method takes a number of optional arguments
allowing you to refine this interaction.

\section{Robot Operating System (ROS)}

Coming soon...